Fission product decay chain equilibria in \zirconia

### 2019-01-25 manuscript.tex
Add iodine explanation

Tellurium on an oxygen site adopts a 0 charge state, which while not obvious (it may be expected to act like oxygen, also a group 6 element), can be reasoned when considering chemical behaviour. Oxygen on its native site will readily adopt a charge state of -2. Tellurium on the other hand, exhibits more metallic behaviour, preferring to get oxidised itself. However, the \ch{V_{O}^{**}} site is highly reducing, and will counter the tellurium ion's proclivity to oxidise. The overall effect on tellurium in this case is to maintain its original electronic structure. Finally, when the tellurium concentration is increased to $10^{-3}$ parts/fu, greater occupancy on the zirconium site is predicted above oxygen pressures of $10^{-17}$ atm, forming the defect \ch{Te_{Zr}^{'''}}. This corresponds to a tellurium charge state of +1. While we would expect an even greater charge state for tellurium, in this case it is the dominant negative defect by an order of magnitude and is necessary to charge compensate positive hole and \ch{Te_{O}^{**}} defects. \\

@@

@@

Brouwer diagrams for xenon defects are shown in Figure \ref{figure:xenonbrouwer}. At a xenon concentration of $10^{-5}$ parts per f.u. \zirconia , the dominant xenon defect is \ch{Xe_{O}^{**}} at low oxygen pressures, and a combination of \ch{Xe_{O}^{**}}, \ch{Xe_{Zr}^{''''}} and \ch{Xe_{Zr}^{'''}} at stoichiometry ($10^{-9}$ atm) and higher. At a higher xenon concentration of $10^{-3}$ parts per f.u. \zirconia , \ch{Xe_{O}^{**}} becomes the dominant defect over a greater oxygen pressure range, from $10^{-30}$ to $10^{-5}$ atm.

@@

        \addplot[no marks, draw=black!70!green] table [x=pO2, y=VO{2},]{dat/xe_tet_10-5.dat}; \addlegendentry{\ch{V_{O}^{\textperiodcentered\textperiodcentered}}}; %\node at (-26.7,-3.3) {\ch{V_{O}^{\textperiodcentered\textperiodcentered}}};

@@

        \addplot[no marks, draw=blue!70!black] table [x=pO2, y=electrons,]{dat/xe_tet_10-3.dat}; %\node at (-27,-1.7) {\ch{e^{'}}};

@@

        \addplot[no marks, dashed, draw=orange] table [x=pO2, y=XesubO{2},]{dat/xe_tet_10-3.dat}; 

@@

        \addplot[no marks, dashed, draw=blue] table [x=pO2, y=XesubZr{-4},]{dat/xe_tet_10-3.dat}; 

@@

        \addplot[no marks, draw=red!85!black] table [x=pO2, y=holes,]{dat/cs_tet_10-3.dat}; %\node at (-2.5,-2.1) {\ch{h^{\textperiodcentered}}};
        \addplot[no marks, draw=black!70!green] table [x=pO2, y=VO{2},]{dat/cs_tet_10-3.dat}; 

@@

        \addplot[no marks, dashed, draw=orange!80!black] table [x=pO2, y=CssubZr{-3},]{dat/cs_tet_10-3.dat}; 

>>>>

Tellurium on an oxygen site adopts a 0 charge state, which while not obvious (it may be expected to act like oxygen, also a group 6 element), can be reasoned when considering chemical behaviour. Oxygen on its native site will readily adopt a charge state of -2. Tellurium on the other hand, exhibits more metallic behaviour, preferring to get oxidised itself. However, the \ch{V_{O}^{**}} site is highly reducing, and will counter the tellurium ion's proclivity to oxidise. The overall effect on tellurium in this case is to maintain its original electronic structure. Finally, when the tellurium concentration is increased to $10^{-3}$ parts/fu, greater occupancy on the zirconium site is predicted above oxygen pressures of $10^{-17}$ atm, forming the defect \ch{Te_{Zr}^{'''}}. This corresponds to a tellurium charge state of +1. While we would expect an even more positive charge state for tellurium, in this case it is the dominant negative defect by an order of magnitude and is necessary to charge compensate positive hole and \ch{Te_{O}^{**}} defects. \\

@@

\subsubsection*{Iodine}

Iodine was previously predicted in Brouwer diagrams (Figure \ref{figure:tikzbrouwerconctet} to produce the defects \ch{I_{O}^{*}} and \ch{I_{Zr}^{'''}} at low and high oxygen pressures respectively. Since the extrinsic defect site is similar for tellurium defects over most oxygen pressures, $\beta$- decay of tellurium could have directly produced these iodine defects without requiring a change of site. Thus, \ch{Te_{O}^{**}} would decay and produce \ch{I_{O}^{*}} with a defect volume of 29.90 \r{A}$^{3}$, and \ch{Te_{Zr}^{'''}} would decay and produce \ch{I_{Zr}^{'''}} with a defect volume of 87.07 \r{A}$^{3}$. At the oxygen site, this means a defect volume change from 2.74 to 29.90 \r{A}$^{3}$. This increase is in large part due to the change in charge state from 0 to -1, as the ions are of similar size otherwise. \\

Similarly at the zirconium site, the defect volume changes from 59.93 to 87.07 \r{A}$^{3}$. Both of these are relatively large increases in defect volume and will come with an increase in lattice stress in the immediate vicinity of the defect, especially if highly constrained. Of course, this  process will take place continuously over time as tellurium atoms decay, with half-lives ranging from seconds to hours depending on the isotope. It is of note however, that PCI failures tend to occur after power ramps with a time delay, even up to an hour after the transient. While not conclusive, this may hint at a mechanism involving decay processes.

@@

Brouwer diagrams for xenon defects are shown in Figure \ref{figure:xenonbrouwer}. At a xenon concentration of $10^{-5}$ parts/fu \zirconia , the dominant xenon defect is \ch{Xe_{O}^{**}} at low oxygen pressures, and a combination of \ch{Xe_{O}^{**}}, \ch{Xe_{Zr}^{''''}} and \ch{Xe_{Zr}^{'''}} at stoichiometry ($10^{-9}$ atm) and higher. At a higher xenon concentration of $10^{-3}$ parts/fu \zirconia , \ch{Xe_{O}^{**}} becomes the dominant defect over a greater oxygen pressure range, from $10^{-30}$ to $10^{-5}$ atm, with \ch{Xe_{Zr}^{''''}} also present in high concentrations at oxygen pressures above $10^{-20}$ atm. As a noble gas, it is expected that xenon in a 0 charge state (\ch{Xe_{O}^{**}} and \ch{Xe_{Zr}^{''''}}) will be prevalent since the full electronic shell provides a stable configuration. \\

@@

        \addplot[no marks, draw=black!70!green] table [x=pO2, y=VO{2},]{dat/xe_tet_10-5.dat}; \addlegendentry{\ch{V_{O}^{\textperiodcentered\textperiodcentered}}}; \node at (-26.7,-3.3) {\ch{V_{O}^{\textperiodcentered\textperiodcentered}}};

@@

        \addplot[no marks, draw=blue!70!black] table [x=pO2, y=electrons,]{dat/xe_tet_10-3.dat}; \node at (-27,-1.7) {\ch{e^{'}}};

@@

        \addplot[no marks, dashed, draw=orange] table [x=pO2, y=XesubO{2},]{dat/xe_tet_10-3.dat}; \node at (-12,-2.7) {\ch{Xe_{O}^{\textperiodcentered\textperiodcentered}}};

@@

        \addplot[no marks, dashed, draw=blue] table [x=pO2, y=XesubZr{-4},]{dat/xe_tet_10-3.dat}; \node at (-24.5,-6) {\ch{Xe_{Zr}^{''''}}};

@@

        \addplot[no marks, draw=red!85!black] table [x=pO2, y=holes,]{dat/cs_tet_10-3.dat}; \node at (-2.5,-2.1) {\ch{h^{\textperiodcentered}}};
        \addplot[no marks, draw=black!70!green] table [x=pO2, y=VO{2},]{dat/cs_tet_10-3.dat}; \node at (-17,-2.2) {\ch{V_{O}^{\textperiodcentered\textperiodcentered}}};

@@

        \addplot[no marks, dashed, draw=orange!80!black] table [x=pO2, y=CssubZr{-3},]{dat/cs_tet_10-3.dat}; \node at (-17,-3.6) {\ch{Cs_{Zr}^{'''}}};
